---
title: "ESD R Package Dependencies"
author: "Andrew M. Dolman"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

This analysis relies on the packages being installed somewhere on the system running it. It will not find reverse dependencies for packages that are not installed.  

This also misses published code, e.g. scripts as supplements to papers, that rely on our R packages. 

```{r}
library(tidyverse)
library(itdepends)
```

```{r}
# ESD public packages 
esd.public <- sort(c("sedproxy", "psem", "PaleoSpec", "orbitalforcing", "ecustools",
                "piccr", "cptPicarr", "TrenchR", "corit", "climproxyrecords"))
```

## List reverse dependencies for all ESD public packages
```{r}
rev.deps <- sapply(esd.public, function(i) sort(tools::dependsOnPkgs(i)))

rev.deps

#plyr::ldply(rev.deps, function(x) data.frame(x, stringsAsFactors = F))
```

## For all ESD public packages, list their dependencies on other ESD packages 
```{r}
ESDDependsOnESD <- function(pkg, pkgs, db){
  x <- tools::package_dependencies(package=pkg, recursive=FALSE, db=db)
  x[[1]][x[[1]] %in% pkgs]
}

IP <- installed.packages()

ESDDependsOnESD("psem", esd.public, db = IP)

sapply(esd.public, function(i) ESDDependsOnESD(i, esd.public, db=IP))
```

# In more detail

Use package "itdepends" to count function call per package

```{r}
all.calls <- data.frame(focal.package = esd.public, stringsAsFactors = FALSE) %>% 
  group_by(focal.package) %>% 
  do({
    itdepends::dep_usage_pkg(.$focal.package)
  })

# exclude "base" and "self"
all.calls.pkg <- all.calls %>% 
  filter(pkg != "base",
         focal.package != pkg) %>% 
  count(pkg) %>% 
  arrange(focal.package, desc(n))

all.calls.pkg
```


```{r}
all.calls.fun <- all.calls %>% 
  filter(pkg != "base", 
         focal.package != pkg) %>% 
  group_by(focal.package, pkg) %>%
  count(fun) %>%
  arrange(focal.package, desc(n)) 
```

### Plot calls per package

```{r}
all.calls.pkg %>% 
  ggplot(aes(x = reorder(pkg, n), y = n, fill = pkg)) +
  geom_col() +
  facet_wrap(~focal.package, scales = "free") +
  coord_flip()
```


### Plot calls per function

```{r}
all.calls.fun %>% 
  ggplot(aes(x = fun, y = n, fill = pkg)) +
  geom_col() +
  facet_wrap(~focal.package, scales = "free") +
  coord_flip()
```

### Just look at psem

```{r}
all.calls.fun %>% 
  filter(focal.package == "psem") %>% 
  ggplot(aes(x = reorder(fun, -n), y = n, fill = pkg)) +
  geom_col() +
  facet_wrap(~focal.package, scales = "free") +
  coord_flip()
```

### Just look at cptPicarr
```{r}
all.calls.fun %>% 
  filter(focal.package == "cptPicarr") %>% 
  ggplot(aes(x = reorder(fun, -n), y = n, fill = pkg)) +
  geom_col() +
  facet_wrap(~focal.package, scales = "free") +
  coord_flip()
```


### Just look at ESD packages

```{r}
all.calls.fun %>% 
  filter(pkg %in% esd.public) %>% 
  ggplot(aes(x = reorder(fun, -n), y = n, fill = pkg)) +
  geom_col() +
  facet_wrap(~focal.package, scales = "free") +
  coord_flip()
```


